Demo arXiv template

### 2022-10-11 paper/paper.qmd
refine on why NDVI is not an index + text editing

Quantities that can't be directly measured is ubiquitous in our society and they can often be relevant for policy making, resource allocation, early prevention and operational purpose.
For example, fire authority would be interested to quantify fire risk since bushfire can have a huge impact on monetary loss, health, and local ecosystem among others.
Climatologists would be interested in monitoring the change in global condition since variability in atmospheric and oceanic condition has a direct impact on the global weather and climate.
Usually more than one variable is related to the concept of interest and they often needs to be integrated to make decisions.
<!-- From a modelling perspective, these unobserved quantities are called latent variables and existing methods can incorporate these latent variables in the model structure.  -->
A commonly used approach here is to construct an index with the observable variables.
This allows analysts to compare the quantity of interest across countries and also cross time.

@@

The meaning of indices varies across disciplines and in this article, an index is defined as a tool to quantify a concept of interest that does not have direct measure.
The concept of interest doesn't have direct measure can because it is impractical to measure the concept of interest at the population level. 
For example, it would be nearly impossible to include all the available stocks in the market to characterise stock market behavior, indices like Dow Jones Industrial Average, S&P 500, and Nasdaq Composite select a representative set of stocks to measure the overall market behavior. 
Also belong to this category is the economic indices like Consumer Price Index, where price changes of a basket of items are weighted to measure inflation. 
The lack of direct measure could also due to the absence of a universal definition on the concept of interest. 
Many natural and social concepts are fundamentally multi-facets and some indices combines measures from different domains to characterise the concept of interest. 
This includes drought indices constructed from meteorological, agricultural, hydrological, and social-economic variables [give some names?]. 
Social development indices like Human Development Index [ref] and Global Liveability Index [ref] measure various aspects on the quality of human capital and urban life.

Despite many measurements have index in its name, they cannot be technically classified as indices according to the definition given above. 
A major reason for these measures to lose its index membership is that the underlying concept of interest is straight forward to measure once defined. 
Sometime variables like precipitation of driest month or percentage of days when maximum temperature is below a certain percentile are classified as climate indices, however, precipitation and temperature can be directly measured by rain gauge and thermometer. 
These variables are transformation on precipitation and temperature. 
They can be used to construct indices but are not indices themselves. 
What is also not indices are many remote sensing based drought measures as they are transformation on one or more remote sensor bands. 
This includes Normalized Difference Vegetation Index (NDVI) [ref], a transformation difference over sum on near-infrared (NIR) and red visible channel, and Vegetation Condition Index [ref], a rescale of NDVI into [0, 1], among others.

TODO: confirm the actual climate indices are indices by our definition
Questions: need reference for stock market indices, CPI?

@@

>>>>

Many variables relevant to decision making cannot be directly measured, however, they are crucial for resource allocation, early prevention, and other operational purpose.
For example, fire authorities would be interested to quantify fire risk since bushfires can have a huge impact on monetary loss, health, and the local ecosystem among others.
Climatologists would be interested in monitoring the change in global conditions since variability in atmospheric and oceanic conditions has a direct impact on global weather and climate.
Usually, more than one variable is related to the concept of interest and they often need to be integrated to make decisions.
<!-- From a modelling perspective, these unobserved quantities are called latent variables and existing methods can incorporate these latent variables in the model structure.  -->
In many domains, constructing an index is a common approach to characterise these unobservable quantities.
The index approach allows researchers to compare the quantity of interest across countries and also cross time.

@@

The meaning of indices varies across disciplines and in this article, an index is defined as a tool to quantify a concept of interest that does not have a direct measure.
The concept of interest doesn't have a direct measure can because it is impractical to measure the concept of interest at the population level. 
For example, it would be nearly impossible to include all the available stocks in the market to characterise stock market behavior, indices like Dow Jones Industrial Average, S&P 500, and Nasdaq Composite select a representative set of stocks to measure the overall market behavior. 
Also belonging to this category are the economic indices like the Consumer Price Index, where price changes of a basket of items are weighted to measure inflation. 
The lack of direct measure could also be due to the absence of a universal definition on the concept of interest. 
Many natural and social concepts are fundamentally multi-facets and some indices combine measures from different domains to characterise the concept of interest. 
This includes drought indices constructed from meteorological, agricultural, hydrological, and social-economic variables [give some names?]. 
Social development indices like Human Development Index [@hdi] and Global Liveability Index [@gli] measure various aspects of the quality of human capital and urban life.

Despite many measurements having the term *index* in their name, they cannot be technically classified as indices according to the definition given above. 
A major reason for these measures to lose their index memberships is that the underlying concept of interest is straightforward to measure once defined. 
Sometimes variables like precipitation of the driest month or percentage of days when maximum temperature is below a certain percentile are classified as climate indices, however, precipitation and temperature can be directly measured by rain gauge and thermometer. 
These variables are transformation on precipitation and temperature. 
They can be used to construct indices but are not indices themselves. 
Similarly, a set of remote sensing indices should not be classified as indices, but variables, since they are measures of electromagnetic wave reflectance.
This includes Normalized Difference Vegetation Index (NDVI) [@tucker_red_1979], derived from the ratio of difference over sum on two segments in the spectrum, also called band: near-infrared (NIR) and red.
So are the "indices" derived from NDVI, e.g. Vegetation Condition Index [@kogan_application_1995]. 
Notice that this does not exclude all the construction derived from remotes sensor variables to be valid indices.
For example, Vegetation Drought Response Index [@brown_vegetation_2008] is a valid index since it integrates climate, satellite, and biophysical variables to quantify vegetation stress.

TODO: clean up terminology:

  - "index" only for valid indices
  - "variable" the most generic dump????
  - "measure" for item quantified by instruments or devices in standard units

FYI:

  - Vegetation Condition Index  = (NDVI - NDVI_min)/ (NDVI_max - NDVI_min)
  - A measure is a variable, the variable itself is not an index. 
  - Not all multivariate "indices" are indices (i.e. those remote sensor transformations), but generally they are. It is possible that an index is constructed from a single variable (i.e. SPI), but be careful on whether it is variable or index.

@@

A conversion of non-normal variables into (approximately) normal.

### 2022-10-11 paper/paper.tex
refine on why NDVI is not an index + text editing

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners, breakable, boxrule=0pt, interior hidden, frame hidden]}{\end{tcolorbox}}\fi

\begin{itemize}
\tightlist
\item
  indices, why do we need them, decision making
\item
  define what is an index, what is not even if they are called index
\item
  only very brief on indices in the introduction
\item
  potential audience + what do you think they should learn from this
  work.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  decision making: need to consider multi-facet for allocating
  resources, {[}\ldots{]}
\item
  to include or constructed to quantify a concept of interest
\end{itemize}

Integrate multiple information to produce a single number summary
(across time).

@@

Indices quantify a concept of interest with no direct measure across
time. The concept of interest doesn't have direct measure can because it
is impractical to measure the concept of interest at the population
level. For example, it would be nearly impossible to include all the
available stocks in the market to characterise stock market behavior,
indices like Dow Jones Industrial Average, S\&P 500, and Nasdaq
Composite select a representative set of stocks to measure the overall
market behavior. Also belong to this category is the economic indices
like Consumer Price Index, where price changes of a basket of items are
weighted to measure inflation. The lack of direct measure could also due
to the absence of a universal definition on the concept of interest.
Many natural and social concepts are fundamentally multi-facets and some
indices combines measures from different domains to characterise the
concept of interest. This includes drought indices constructed from
meteorological, agricultural, hydrological, and social-economic
variables {[}give some names?{]}. Social development indices like Human
Development Index {[}ref{]} and Global Liveability Index {[}ref{]}
measure various aspects on the quality of human capital and urban life.

Despite many measurements have index in its name, they cannot be
technically classified as indices according to the definition given
above. A major reason for these measures to lose its index membership is
that the underlying concept of interest is straight forward to measure
once defined. Sometime variables like precipitation of driest month or
percentage of days when maximum temperature is below a certain
percentile are classified as climate indices, however, precipitation and
temperature can be directly measured by rain gauge and thermometer.
These variables are transformation on precipitation and temperature.
They can be used to construct indices but are not indices themselves.
What is also not indices are many remote sensing based drought measures
as they are transformation on one or more remote sensor bands. This
includes Normalized Difference Vegetation Index (NDVI) {[}ref{]}, a
transformation difference over sum on near-infrared (NIR) and red
visible channel, and Vegetation Condition Index {[}ref{]}, a rescale of
NDVI into {[}0, 1{]}, among others.

TODO: confirm the actual climate indices are indices by our definition
Questions: need reference for stock market indices, CPI?

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-buja_elements_1988}{}}%
Buja, A, D Asimov, C Hurley, and JA McDonald. 1988. {``Elements of a
Viewing Pipeline for Data Analysis.''} In \emph{Dynamic Graphics for
Statistics}, 277--308. Wadsworth, Belmont.

@@

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, frame hidden, enhanced, boxrule=0pt, breakable, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

\newpage

@@

Many variables relevant to decision making cannot be directly measured,
however, they are crucial for resource allocation, early prevention, and
other operational purpose. For example, fire authorities would be
interested to quantify fire risk since bushfires can have a huge impact
on monetary loss, health, and the local ecosystem among others.
Climatologists would be interested in monitoring the change in global
conditions since variability in atmospheric and oceanic conditions has a
direct impact on global weather and climate. Usually, more than one
variable is related to the concept of interest and they often need to be
integrated to make decisions. In many domains, constructing an index is
a common approach to characterise these unobservable quantities. The
index approach allows researchers to compare the quantity of interest
across countries and also cross time.

@@

The meaning of indices varies across disciplines and in this article, an
index is defined as a tool to quantify a concept of interest that does
not have a direct measure. The concept of interest doesn't have a direct
measure can because it is impractical to measure the concept of interest
at the population level. For example, it would be nearly impossible to
include all the available stocks in the market to characterise stock
market behavior, indices like Dow Jones Industrial Average, S\&P 500,
and Nasdaq Composite select a representative set of stocks to measure
the overall market behavior. Also belonging to this category are the
economic indices like the Consumer Price Index, where price changes of a
basket of items are weighted to measure inflation. The lack of direct
measure could also be due to the absence of a universal definition on
the concept of interest. Many natural and social concepts are
fundamentally multi-facets and some indices combine measures from
different domains to characterise the concept of interest. This includes
drought indices constructed from meteorological, agricultural,
hydrological, and social-economic variables {[}give some names?{]}.
Social development indices like Human Development Index (United Nations
Development Programme 2022) and Global Liveability Index (Economist
Intelligence Unit 2019) measure various aspects of the quality of human
capital and urban life.

Despite many measurements having the term \emph{index} in their name,
they cannot be technically classified as indices according to the
definition given above. A major reason for these measures to lose their
index memberships is that the underlying concept of interest is
straightforward to measure once defined. Sometimes variables like
precipitation of the driest month or percentage of days when maximum
temperature is below a certain percentile are classified as climate
indices, however, precipitation and temperature can be directly measured
by rain gauge and thermometer. These variables are transformation on
precipitation and temperature. They can be used to construct indices but
are not indices themselves. Similarly, a set of remote sensing indices
should not be classified as indices, but variables, since they are
measures of electromagnetic wave reflectance. This includes Normalized
Difference Vegetation Index (NDVI) (Tucker 1979), derived from the ratio
of difference over sum on two segments in the spectrum, also called
band: near-infrared (NIR) and red. So are the ``indices'' derived from
NDVI, e.g.~Vegetation Condition Index (Kogan 1995). Notice that this
does not exclude all the construction derived from remotes sensor
variables to be valid indices. For example, Vegetation Drought Response
Index (Brown et al. 2008) is a valid index since it integrates climate,
satellite, and biophysical variables to quantify vegetation stress.

TODO: clean up terminology:

\begin{itemize}
\tightlist
\item
  ``index'' only for valid indices
\item
  ``variable'' the most generic dump????
\item
  ``measure'' for item quantified by instruments or devices in standard
  units
\end{itemize}

FYI:

\begin{itemize}
\tightlist
\item
  Vegetation Condition Index = (NDVI - NDVI\_min)/ (NDVI\_max -
  NDVI\_min)
\item
  A measure is a variable, the variable itself is not an index.
\item
  Not all multivariate ``indices'' are indices (i.e.~those remote sensor
  transformations), but generally they are. It is possible that an index
  is constructed from a single variable (i.e.~SPI), but be careful on
  whether it is variable or index.
\end{itemize}

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-brown_vegetation_2008}{}}%
Brown, Jesslyn F., Brian D. Wardlow, Tsegaye Tadesse, Michael J. Hayes,
and Bradley C. Reed. 2008. {``The {Vegetation} {Drought} {Response}
{Index} ({VegDRI}): {A} {New} {Integrated} {Approach} for {Monitoring}
{Drought} {Stress} in {Vegetation}.''} \emph{GIScience \& Remote
Sensing} 45 (1): 16--46.
\url{https://doi.org/10.2747/1548-1603.45.1.16}.

\leavevmode\vadjust pre{\hypertarget{ref-buja_elements_1988}{}}%
Buja, A, D Asimov, C Hurley, and JA McDonald. 1988. {``Elements of a
Viewing Pipeline for Data Analysis.''} In \emph{Dynamic Graphics for
Statistics}, 277--308. Wadsworth, Belmont.

\leavevmode\vadjust pre{\hypertarget{ref-gli}{}}%
Economist Intelligence Unit. 2019. {``The Global Liveability Index
2019.''} The Economist.
\url{https://www.cbeinternational.ca/pdf/Liveability-Free-report-2019.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-kogan_application_1995}{}}%
Kogan, F. N. 1995. {``Application of Vegetation Index and Brightness
Temperature for Drought Detection.''} \emph{Advances in Space Research},
Natural {Hazards}: {Monitoring} and {Assessment} {Using} {Remote}
{Sensing} {Technique}, 15 (11): 91--100.
\url{https://doi.org/10.1016/0273-1177(95)00079-T}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-tucker_red_1979}{}}%
Tucker, Compton J. 1979. {``Red and Photographic Infrared Linear
Combinations for Monitoring Vegetation.''} \emph{Remote Sensing of
Environment} 8 (2): 127--50.
\url{https://doi.org/10.1016/0034-4257(79)90013-0}.

\leavevmode\vadjust pre{\hypertarget{ref-hdi}{}}%
United Nations Development Programme. 2022. {``Human Development Report
2021-22.''} New York. \url{http://report.hdr.undp.org}.
